---
title: "HR2_personnel"
output:
  html_document: default
  pdf_document: default
---



```{r}

library(tidyverse)
#library(httr)
#library(jsonlite)
library(lubridate)
library(readr)   
library(dplyr)
#library(psych)
#library(dlookr)
library(Hmisc)
library(corrplot)
```


```{r setup, include=FALSE}

rawdata <- "C:/Users/small/Documents/Mongomery_College/2021_fall/DATA_205/datasets"
setwd(rawdata)
knitr::opts_knit$set(root.dir = rawdata)
getwd()
  
```


```{r}

personnel <- read.csv("PMR_MLS__Management__Report.csv", header = TRUE, fileEncoding="UTF-8-BOM")

```


```{r}
str(personnel)

```

```{r}
library(janitor)

personnel1 <- personnel %>%
  clean_names()
personnel1
```





```{r}
personnel1$gender <- as.factor(personnel1$gender) 
personnel1$length_of_service <- as.numeric(personnel1$length_of_service) 
personnel1$age <- as.numeric(personnel1$age) 

class(personnel1$gender)
class(personnel1$length_of_service)
class(personnel1$age)

```

```{r}
str(personnel1)

```

```{r}
library(dlookr)
personnel1 %>%
group_by(gender) %>% 
describe(length_of_service, age) 

```


```{r}
personnel1 %>%
group_by(gender) %>% 
normality(length_of_service, age)

```


```{r}
personnel1 %>%
  normality(length_of_service, age) %>%
  filter(p_value <= 0.01)
```

```{r}
personnel1 %>%
plot_normality(length_of_service, age)
```





```{r}
salary = c("80-89K", "90-99K", "100-109K", "110-119K", "120-129K", "130-139K" , "140-149K" , ">=150K")

personnel1$salary_range <- factor(personnel1$salary_range, levels=salary, ordered = TRUE)

personnel1$salary_range


#personnel1$salary_range <- as.factor(personnel1$salary_range) 

class(personnel1$salary_range)
```


```{r}
table(personnel1$salary_range)
```


```{r}
type_counts <- table(personnel1$salary_range)
type_counts / sum(type_counts)
```


```{r}

target_salary <- target_by(personnel1, salary_range)

```


```{r}
salary_gender <- relate(target_salary, generation)
salary_gender
summary(salary_gender)

```



```{r}
salary_age <- relate(target_salary, age)
salary_age
summary(salary_age)
```


```{r}
plot(salary_age)
```

```{r}
salary_generation <- relate(target_salary, generation)
salary_generation
summary(salary_generation)

```

According to these results, generation does contribute to salary range (p-value < 0.05%)



```{r}
personnel1$ethnic_origin[personnel1$ethnic_origin == "White (Not Hispanic or Latino)"] <- "White"
personnel1$ethnic_origin[personnel1$ethnic_origin== "Asian (Not Hispanic or Latino)"] <- "Asian"
personnel1$ethnic_origin[personnel1$ethnic_origin == "Black or African American (Not Hispanic or Latino)"] <- "Black or African American"
personnel1$ethnic_origin[personnel1$ethnic_origin == "American Indian or Alaska Native (Not Hispanic or Latino)"] <- "American Indian or Alaska Native"
personnel1$ethnic_origin[personnel1$ethnic_origin == "Two or More Races (Not Hispanic or Latino)"] <- "Two or More Races"


```



```{r}
salary_ethnic_origin <- relate(target_salary, ethnic_origin)
salary_ethnic_origin
summary(salary_ethnic_origin)
```

According to these results, ethnicity does not contribute to salary range (p-value > 0.05%)




```{r}
salary_length_of_service <- relate(target_salary, length_of_service)
salary_length_of_service
summary(salary_length_of_service)
```

```{r}
plot(salary_length_of_service)
```

```{r}
# Are salary and gender related?
# Null Hypothesis: salary and gender are independent

sal_gender <- table(personnel1$salary_range, personnel1$ethnic_origin)
sal_gender
chisq.test(sal_gender)
```




```{r}
# Are salary and length_of_service related?
# Null Hypothesis: salary and length_of_service are independent
sal_years <- table(personnel1$salary_range, personnel1$length_of_service)
sal_years
chisq.test(sal_years)
```

```{r}
# Are salary and age related?
# Null Hypothesis: salary and age are independent
sal_age <- table(personnel1$age, personnel1$age)
sal_age
chisq.test(sal_age)
```

```{r}
# Are salary and ethnicity related?
# Null Hypothesis: salary and ethnicity are independent
sal_ethnicity <- table(personnel1$salary_range, personnel1$ethnic_origin)
sal_ethnicity
chisq.test(sal_ethnicity)
```

```{r}
plot1 <- personnel1 %>%
  ggplot(aes(x = salary_range, fill = ethnic_origin)) +
  geom_histogram(stat="count", binwidth = 5, color = "white")+
  scale_fill_discrete(name = "Ethnicity", labels = c("American Indian or Alaska Native", "Asian","Black or African American", "Hispanic or Latino", "Two or More Races", "Unreported", "White")) +
  labs(x = "Salary (in dollars)", y = "Frequency", 
       title = "MoCo Employees Salary Distribution by Ethnicity")+
  theme_minimal() +
  scale_fill_brewer(palette="Set1")
plot1
```


```{r}
personnel1$generation[personnel1$generation == "Millennial Generation"] <- "Millennial"
personnel1$generation[personnel1$generation == "Traditionalist/Silent Generation"] <- "Silent"


```



```{r}
plot1 <- personnel1 %>%
  ggplot(aes(x = salary_range, fill = grade)) +
  geom_histogram(stat="count", binwidth = 5, color = "white")+
  scale_fill_discrete(name = "Ethnicity", labels = c("M1", "M2", "M3")) +
  labs(x = "Salary (in dollars)", y = "Frequency", 
       title = "MoCo Employees Salary Distribution by Grade")+
  theme_minimal() +
  scale_fill_brewer(palette="Set1")
plot1
```



```{r}
ggplot(data = personnel1) +
  geom_bar(mapping = aes(x = salary_range,
                       fill = gender)) +
  labs(x = "Salary (in dollars)",
       title = "MoCo Employees Salary Distribution by Gender")+
  theme_minimal() +
  scale_fill_brewer(palette="Set1")
  
```

{r fig.width = 7, fig.height = 4}


```{r}
table_gender_salary <- personnel1 %>%
  dplyr::count(grade, salary_range)

ggplot(data = table_gender_salary) +
  geom_tile(mapping = aes(x = grade, 
                          y = salary_range, fill = n)) +
  scale_fill_gradientn(colors = brewer.pal(5, "RdYlGn")) +
labs(x = "Grade", y = "Salary (in dollars)", 
       title = "MoCo Employees Salary Distribution by Grade") +
  theme_minimal()
 # facet_grid(.~ grade)
 
```


Is there a relationship between gender and salary range?

This is going to be a Chi-Square test. We will test using α=0.05.
Ho: gender and salary range are independent.
Ha: gender and salary range are dependent.



```{r}
gender_salary <- table(personnel1$grade, 
                            personnel1$grade)

result5 <- chisq.test(gender_salary)
result5
```

P-value: 0.4017 > 0.05 = α.

Conclusion: Fail to reject Ho.

```{r}
result3$residuals

```

```{r fig.width = 7, fig.height = 8}
corrplot(result3$residuals, is.corr = FALSE)
```









