---
title: "HR2_salaries"
output: html_document
---



```{r}

library(tidyverse)
#library(httr)
#library(jsonlite)
#library(lubridate)
library(readr)   
library(dplyr)
#library(testthat)
#library(psych)
library(dlookr)
#library(Hmisc)
library(corrplot)

```




```{r setup, include=FALSE}

rawdata <- "C:/Users/small/Documents/Mongomery_College/2021_fall/DATA_205/datasets"
setwd(rawdata)
knitr::opts_knit$set(root.dir = rawdata)
getwd()
  
```


```{r}

salaries <- read.csv("Employee_Salaries_-_2020.csv", header = TRUE, fileEncoding="UTF-8-BOM")


```



```{r}
str(salaries)
```

```{r}
library(janitor)

salaries1 <- salaries %>%
  clean_names()
salaries1
```



```{r}
salaries1 %>%
group_by(gender) %>% 
describe(base_salary)
```




```{r}
salaries1 %>%
#group_by(department_name) %>% 
describe(base_salary)

```



```{r}

salaries1 %>%
group_by(gender) %>% 
normality(base_salary)

```


```{r}
salaries1 %>%
  mutate(log_base_salary = log(base_salary)) %>%
  group_by(department_name) %>%
  normality(log_base_salary)
 # filter(p_value > 0.01)

```

```{r}
salaries1 %>%
  mutate(log_base_salary = log(base_salary)) %>%
  group_by(gender) %>%
  normality(log_base_salary)

```


```{r}
library(dlookr)
plot_normality(salaries1)
```


```{r}
correlate(salaries1)
```


```{r}
plot_correlate(salaries1)
```

```{r}
salaries1$gender <- as.factor(salaries1$gender) 

class(salaries1$gender)
```

```{r}
target_base_salary<- target_by(salaries1, base_salary)

cat_categor <- relate(target_base_salary, gender)
cat_categor

summary(cat_categor)
```

According to these results, gender does contribute to salary amount (p-value < 0.05%).

Although the coefficient for the predictor sex is found to be significant in the model, sex only explains around 0.3% of the variance in salary in this model. 



```{r}
plot(cat_categor)
```

```{r}
plot1 <- salaries1 %>%
   ggplot(aes(gender, base_salary, fill = gender)) + 
  ggtitle("Distribution of MoCo Employees Salary by Gender") +
  xlab("Gender") +
  ylab("Salary") +
  geom_boxplot() +
  scale_fill_discrete(name = "Gender", labels = c("Female", "Male")) +
  theme_minimal() +
  scale_fill_brewer(palette="Set1")
plot1
```

It appears average of salaries for two gender is slightly different. Is this actually true or could it have just happened by chance for our sample of data?


This is going to be an ANOVA test and we will use α=0.05.

Ho:μbx=μbr=μm=μq=μsi (In other words, there is no relationship between salary and gender)

Ha: Not all μ are equal. (There is a relationship between salary and gender)


```{r}
results <- aov(salaries1$base_salary ~ salaries1$gender)

summary(results)
```
P-value: 0.0000000282 < 0.05=α.

Conclusion: Reject Ho.



It appears that women has a lower salary than men. Is this true? Or did this just happen by chance with our sample?

This is going to be a two sample t-test. We will test at the 95% confidence level.

Ho:μm=μb (In other words, the salaries are the same)
Ha:μb<μm (In other words, the salaries of women are less then salaries of men)

```{r}

women <- salaries1 %>%
  filter(gender == "F")

men <- salaries1 %>%
  filter(gender == "M")

t.test(women$base_salary, men$base_salary, alternative="less",
       conf.level = 0.95)

```

P-value: 0.00000001857 < 0.05 = α.

Conclusion: Reject Ho.



What about the distribution salary in the Departments with biggest and smallest salaries?


```{r}
salaries_big_sal <- salaries1 %>%
  select(base_salary, department_name) %>%
  group_by(department_name) %>%
#arrange(-base_salary) 
 arrange(base_salary) 
salaries_big_sal

```

For department with the buggiest salaries - 


Ho:μm=μb (In other words, the salaries are the same)
Ha:μb<μm (In other words, the salaries of women are less then salaries of men)


```{r}

women1 <- salaries1 %>%
  filter(gender == "F" & department_name == "Offices of the County Executive")

men1 <- salaries1 %>%
  filter(gender == "M" & department_name == "Offices of the County Executive")

t.test(women1$base_salary, men1$base_salary, alternative="less",
       conf.level = 0.95)


```

P-value: 0.001843 < 0.05 = α.

Conclusion: Reject Ho.



Ho:μm=μb (In other words, the salaries are the same)
Ha:μb<μm (In other words, the salaries of women are less then salaries of men)


```{r}

women2 <- salaries1 %>%
  filter(gender == "F" & department_name == "Department of Police")

men2 <- salaries1 %>%
  filter(gender == "M" & department_name == "Department of Police")

t.test(women2$base_salary, men2$base_salary, alternative="less",
       conf.level = 0.95)

```

P-value: 0.00000000000000022 < 0.05 = α.

Conclusion: Reject Ho.


Ho:μm=μb (In other words, the salaries are the same)
Ha:μb<μm (In other words, the salaries of women are less then salaries of men)

```{r}
women3 <- salaries1 %>%
  filter(gender == "F" & department_name == "Department of Health and Human Services")

men3 <- salaries1 %>%
  filter(gender == "M" & department_name == "Department of Health and Human Services")

t.test(women3$base_salary, men3$base_salary, alternative="less",
       conf.level = 0.95)
```

P-value: 0.0000004074 < 0.05 = α.

Conclusion: Reject Ho.


For department with the smallest salaries - 


Ho:μm=μb (In other words, the salaries are the same)
Ha:μb<μm (In other words, the salaries of women are less then salaries of men)


```{r}
women4 <- salaries1 %>%
  filter(gender == "F" & department_name == "Office of Agriculture")

men4 <- salaries1 %>%
  filter(gender == "M" & department_name == "Office of Agriculture")

t.test(women4$base_salary, men4$base_salary, alternative="less",
       conf.level = 0.95)
```


P-value: 0.1891 > 0.05 = α.

Conclusion: Fail to reject Ho.



```{r}
men7 <- salaries1 %>%
  filter(gender == "M" & department_name == "Office of Agriculture")

women7 <- salaries1 %>%
  filter(gender == "F" & department_name == "Office of Agriculture")

t.test(men7$base_salary, women7$base_salary, alternative="less",
       conf.level = 0.95)
```

```{r}
women5 <- salaries1 %>%
  filter(gender == "F" & department_name == "Department of Public Libraries")

men5 <- salaries1 %>%
  filter(gender == "M" & department_name == "Department of Public Libraries")

t.test(women5$base_salary, men5$base_salary, alternative="less",
       conf.level = 0.95)
```

P-value: 0.02504 < 0.05 = α.

Conclusion: Reject Ho.


```{r}
women6 <- salaries1 %>%
  filter(gender == "F" & department_name == "Department of Permitting Services")

men6 <- salaries1 %>%
  filter(gender == "M" & department_name == "Department of Permitting Services")

t.test(women6$base_salary, men6$base_salary, alternative="less",
       conf.level = 0.95)
```

P-value: 0.09165 > 0.05 = α.

Conclusion: Fail to reject Ho.




```{r}
summary(salaries1$base_salary)
```


```{r}
library(plyr)
salaries2 <- ddply(salaries1, "gender", summarise, base_salary.mean=mean(base_salary))
salaries2
```

The salaries are different, but gender variable only explains around 0.3% of the variance.

Which other variables in the dataset affect salary and how much?

Let’s take a brief look at salary regressed upon the variables in simple linear regression models


```{r}
mod_dep <- lm(base_salary~department_name, data = salaries1)
summary(mod_dep)
```
According to these results, the name of department does contribute to salary amount as expected (p-value < 0.05)

The amount of variance in salary explained by gender in this model is around 18.99%


```{r}
mod_grade <- lm(base_salary~grade, data = salaries1)
summary(mod_grade)
```


According to these results, gender does contribute to salary amount as expected (p-value < 0.05)

The amount of variance in salary explained by grade in this model is around 83.99%


Cheacking all variables to collinearity


```{r}
library(DataExplorer)
plot_correlation(salaries1)
```


Build model with all meaningful variables


```{r}
model_full <- lm(base_salary ~ department_name + grade, data = salaries1)
summary(model_full)


```
The combination of four variables (department_name + grade) has the biggest adjusted R-squared value - 87% 




```{r}
plot(model_full)
```



```{r fig.width = 6, fig.height = 4}
salaries1  %>%
group_by(department_name, gender) %>%
summarize(avg = mean(base_salary, na.rm=T)) %>%

ggplot(aes(x = reorder(department_name, +avg), avg, fill = gender)) +
geom_bar(stat = "identity") +
coord_flip() +
theme_minimal() +
  theme(legend.position = "bottom") + #set position of the legend
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Distribution of MoCo Employees Salary by Departments in 2020",
       x = "Deparmnents", 
       y = "Salary (in dollars)",  
       fill = "Gender")



```



```{r}
salaries_m <- salaries1 %>%
  filter(gender == "M")

```



```{r}
salaries_m %>%
group_by(department_name, gender) %>%
summarize(avg = mean(base_salary, na.rm=T)) %>%

ggplot(aes(x = reorder(department_name, +avg), avg, fill = gender)) +

geom_bar(stat = "identity") +

coord_flip() +

theme_minimal()
```



```{r}
salaries_hist1 <- salaries1 %>%
  ggplot(aes(x = base_salary, fill = gender), options(scipen = 999)) +
  geom_histogram(position="dodge", bins = 25)+
 # scale_fill_discrete(name = "Gender", labels = c("Female", "Male"))+
  
  geom_vline(data=salaries2, aes(xintercept=base_salary.mean,  colour = gender),
               linetype="dashed", size=1) +
  
  labs(x = "Salary (in dollars)", y = "Frequency", 
       title = "MoCo Employees Salary Distribution")+
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")

salaries_hist1
```


```{r}
library(treemap)
treemap(salaries1, index="department_name", vSize="base_salary", #set size
        vColor="base_salary", #set color
        type="manual",
        title="Treemap of Salary by Departments", palette="RdYlBu")
        
        
```
